Dependent Object Types

### 2012-09-30 doc/fool/article.tex
spell out glb and lub throughout

\begin{abstract}
We propose a new type-theoretic foundation of Scala and languages like
it: the Dependent Object Types (DOT) calculus. DOT models Scala's
path-dependent types, abstract type members and its mixture of nominal
and structural typing through the use of refinement types. The core
formalism makes no attempt to model inheritance and mixin
composition. DOT normalizes Scala's type system by unifying the
constructs for type members and by providing classical intersection
and union types which simplify glb/lub computations.

@@

This is problematic because glbs and lubs play a central role in
Scala's type inference. For example, in order to infer the type of an
\code{if} expression such as
\begin{lstlisting}
  if (cond) ((a: A) => c: C) else ((b: B) => d: D)
\end{lstlisting}
type inference tries to compute the glb of \code{A} and \code{B} and
the lub of \code{C} and \code{D}. The absence of universal glbs and
lubs makes type inference more brittle and more unpredictable.

@@

\subsection{Glbs and Lubs}

In DOT, the glb of two types is their intersection and their lub,
their union. Recall the introductory example which was problematic in
Scala:
\begin{lstlisting}
  trait A { type T <: A }
  trait B { type T <: B }
\end{lstlisting}

@@

The glb is of $p.A_c$ and $p.B_c$ is $p.A_c \tand p.B_c$ by
definition.

@@

%\section{Examples}
%TODO: WIP
%
%\includegraphics[scale=0.7]{../../src/redex/ex_glb}

>>>>

\begin{abstract}
We propose a new type-theoretic foundation of Scala and languages like
it: the Dependent Object Types (DOT) calculus. DOT models Scala's
path-dependent types, abstract type members and its mixture of nominal
and structural typing through the use of refinement types. The core
formalism makes no attempt to model inheritance and mixin
composition. DOT normalizes Scala's type system by unifying the
constructs for type members and by providing classical intersection
and union types which simplify greatest lower bound and least upper bound computations.

@@

This is problematic because greatest lower bounds and least upper bounds play a central role in
Scala's type inference. For example, in order to infer the type of an
\code{if} expression such as
\begin{lstlisting}
  if (cond) ((a: A) => c: C) else ((b: B) => d: D)
\end{lstlisting}
type inference tries to compute the greatest lower bound of \code{A} and \code{B} and
the least upper bound of \code{C} and \code{D}. The absence of universal greatest lower bounds and
least upper bounds makes type inference more brittle and more unpredictable.

@@

\subsection{Greatest Lower Bounds and Least Upper Bounds}

In DOT, the greatest lower bound of two types is their intersection and their least upper bound,
their union. Recall the introductory example which was problematic in
Scala:
\begin{lstlisting}
  trait A { type T <: A }
  trait B { type T <: B }
\end{lstlisting}

@@

The greatest lower bound of $p.A_c$ and $p.B_c$ is $p.A_c \tand p.B_c$ by
definition.

@@
